---
title: "hw04"
author: "Peter Whitman"
date: "10/9/2017"
output: github_document
---

```{r}
library(gapminder)
library(knitr)
library(tidyverse)
library(reshape2)
library(devtools)
library(readxl)
library(tidyr)
```

# 1. General data reshaping and relationship to aggregation

#### Make a tibble with one row per year and columns for life expectancy for two or more countries.
* Use knitr::kable() to make this table look pretty in your rendered homework.
* Take advantage of this new data shape to scatterplot life expectancy for one country against that of another.

```{r}
df.1 <- gapminder %>% filter(country == "Canada" | country == "United States" | country == "Uganda" | country == "Australia" | country == "Mongolia" | country == "Saudi Arabia") %>% select(year, country, lifeExp) %>% spread(key = country, value = lifeExp) 
kable(df.1)
```

*Above is a tibble with one row per year and columns for life expectancy for two or more countries. The table is made pretty by the use of kable*

```{r}
df.2 <- melt(df.1, id.vars = "year")
kable(df.2)
```

*Here is the same information seen in the first tibble, but outputed in long format. This will allow me to construct a scatterplot with life expectancy (value) on the y axis, year on x axis, and countries represented by color*

```{r}
graph1 <- df.2 %>% ggplot(aes(year, value, color = variable)) + geom_point() + labs(x = "Year", y = "Life Expectancy", title = "Life Expectancy Change Over Time") + scale_colour_discrete("Countries")
graph1
```

*This is a scatterplot that depicts the change in life expectancy for each selected country between 1952 and 2007*

```{r}
graph1 + geom_smooth() 
```

*Here is the same figure, but outputed using geom_smooth, which helps the viewer understand the trends in the data*

# 2. Join, merge, look up

#### Create a second data frame, complementary to Gapminder. Join this with (part of) Gapminder using a  dplyr join function and make some observations about the process and result. Explore the different types of joins. Examples of a second data frame you could build:

```{r}
HDI <- read_excel("HumanDevelopmentIndex.xlsx")
glimpse(HDI)
```

*This provides a brief glimpse of the properties of the data frame that I used for the second task. It is table that depicts the changes in human development index (HDI) between 1990 and 2015 for all UN countries. The table consists of 28 variables and 188 observations*

```{r}
viz1 <- gather(HDI, "year", "HDI", 3:28) %>% select(Country, year, HDI) %>% filter(!is.na(HDI)) %>% group_by(Country)
viz1
```

*In an effort to allow the viewer a better understanding of the dataset, I have reshaped the dataset for easier visualization with ggplot. The above table is in long format*

```{r}
viz1 %>% ggplot(aes(year, HDI)) + geom_line(aes(group=Country, alpha = 4)) + labs(title = "Change in Human Development Index for UN Countries", y= "Human Development Index (HDI)", x = "Year") + scale_x_discrete( breaks=c(1990, 1995, 2000, 2005, 2010, 2015)) + theme(legend.position="none") + geom_hline(yintercept=0.8,
               linetype="dashed", color = "blue", size = 1)
```

*Here I have outputed the HDI data using ggplot. Each line represents one country. The dashed line represents the threshold that the UN uses to dilineate developed countries from developing countries*

```{r}
p1 <- left_join(gapminder, HDI, by = c("country" = "Country"))
p1
```

*Here I have joined the HDI data to the gapminder data by country*

```{r}
dev1 <- p1 %>% filter(year == 2007) %>% select(country, continent, lifeExp, pop, gdpPercap, `2007`) 
dev1
```

*I would like to take a look at the number of countries within each continent that fall above and bellow the development threshold in 2007. So, here I have filtered out all survey years except for 2007 and selected relevant variables*


```{r}
dev2 <- dev1 %>% mutate(dev_threshold=c("developing2007", "developed2007")[(`2007` >= 0.800) + 1]) %>% filter(!is.na(`2007`))
dev2
```

*Here I've created a development index (0.8), which is the threshold that the UN uses to determine whether or not a country is developed. I've also created a new column that lets us know whether or not a country is above or below the threshold in 2007*

```{r}
lm.1 <- lm(dev2$`2007` ~ dev2$lifeExp)
summary(lm.1)
lm.2 <- lm(dev2$`2007` ~ dev2$gdpPercap)
summary(lm.2)
lm.3 <- lm(dev2$`2007` ~ dev2$pop)
summary(lm.3)
```

*Although I can guess, I don't know how HDI was computed. So, I want to take a look at the relationship between HDI and life expectancy, gdp/capita, and population. To do this, I've run a linear regression, which allows me to ouput an r^2^ and adjusted r^2^. Higher r^2^ values indicate a higher correlation between the variables.*

```{r}
dev2 %>% ggplot(aes(gdpPercap, `2007`, color = dev_threshold)) + facet_wrap(~ continent) + geom_point(size = 1) + scale_color_discrete("Development Status") + labs(title = "HDI vs GDP/Capita in 2007 by Continent", x = "GDP/capita", y = "Human Development Index (HDI)") + scale_x_continuous(breaks = c(0, 20000, 40000, 60000))
```

```{r}
dev2 %>% ggplot(aes(lifeExp, `2007`, color = dev_threshold)) + facet_wrap(~ continent) + geom_point(size = 1) + scale_color_discrete("Development Status") + labs(title = "HDI vs Life Expectancy in 2007 by Continent", x = "Life Expectancy", y = "Human Development Index (HDI)") 
```

*I created two clusters of scatterplots that depict HDI vs the variables with acceptable r^2^ values -- life expectancy and GDP/capita*

```{r}
dev3 <- dev2 %>% group_by(continent) %>% count(dev_threshold)
kable(dev3)
```

*Here I am counting up the number of countries that are developed and developing per continent in 2007. This table is in long format*

```{r}
dev4 <- dev3 %>% spread(key = dev_threshold, value = n) 
dev4[is.na(dev4)] <- 0
kable(dev4)
```

*Here I've computed the same data in wide format, which allows me to discern with greater ease the continents that do not have any countries that are developed (Africa) or that don't have any countries that are developed (Oceania)*

```{r}
dev5 <- melt(dev4, id.vars = "continent") 
dev5 %>% 
ggplot(aes(continent, fill = variable, y = value)) +
    geom_col(position = "dodge") + labs(x="Continent", 
          y="Number of Countries",
          title="Number of Developed and Developing Countries by Continent in 2007") 
```

*This bar chart is a depicts the data presented in the table above.*

```{r}
dev6 <- p1 %>% filter(year == 1992) %>% select(country, continent, lifeExp, pop, gdpPercap, `1992`)
dev7 <- dev6 %>% mutate(dev_threshold=c("developing1992", "developed1992")[(`1992` >= 0.800) + 1]) %>% filter(!is.na(`1992`))
dev8 <- dev7 %>% group_by(continent) %>% count(dev_threshold)
dev9 <- dev8 %>% spread(key = dev_threshold, value = n)
dev9[is.na(dev9)] <- 0
kable(dev9)
```

*In an effort to compare the number of countries that have become developed between 1992 and 2007 per continent, I have followed the same workflow as above, but with 1992 instead of 2007*

```{r}
dev10 <- inner_join(dev4, dev9, by = "continent")
kable(dev10)
```

*I then joined the 2007 developed vs developing table to the 1992 developed vs developing table using an inner_join*

```{r}
dev11 <- dev10 %>% select(continent, developed2007, developed1992)
kable(dev11)
```

*The developing columns were removed*

```{r}
dev12 <- melt(dev11, id.vars = "continent") 
dev12 %>% ggplot(aes(continent, fill = variable, y = value)) +
    geom_col(position = "dodge") + labs(x="Continent", 
          y="Number of Countries",
          title="Number of Developed Countries by Continent in 1992 and 2007") + scale_fill_discrete("Developed Countries")
```

*Here is the final graph depicting the change in the number of developed countries between 1992 and 2007*