---
title: "hw05"
author: "Peter Whitman"
date: "10/19/2017"
output: github_document
---

```{r}
library(gapminder)
library(singer)
library(knitr)
library(tidyverse)
library(reshape2)
library(devtools)
library(readxl)
library(tidyr)
library(forcats)
library(ggmap)
library(RColorBrewer)
```

## Factor Management
* Define factor variables

```{r}
data("singer_locations")
glimpse(singer_locations)
```

```{r}
release_fac <- singer_locations %>% 
  transform(release = factor(release))
glimpse(release_fac$release)
fct_count(release_fac$release)
```

```{r}
duration_length <- c("long", "short")
as.factor(duration_length)
```

```{r}
as_factor(duration_length)
```

```{r}
factor(duration_length)
```

* Drop factor / levels

```{r}
dim(release_fac)
nlevels(release_fac$release)
```

```{r}
release_drop <- release_fac %>%
  filter(year != 0) %>%
  droplevels()
dim(release_drop)
```

```{r}
nlevels(release_drop$release)
```

*100 records were filtered out from the dataset, which corresponded to a reduction of 93 levels*

* Reorder levels based on knowledge from data.

```{r}
reorder <- levels(fct_reorder(release_drop$release, release_drop$artist_familiarity, max,desc = FALSE))
head(reorder)
```

```{r}
arrange1 <- release_drop %>%
  filter(year >= 1995 & city == "Minneapolis, MN") %>%
  arrange(artist_familiarity) %>% 
  ggplot(aes(release, artist_familiarity)) + geom_col() + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + labs(title = "Album Releases in Minneapolis by Artist Familiarity (1995 - Present)", x = "Album Title", y = "Artist Familiarity")
arrange1
```

```{r}
arrange2 <- release_drop %>% 
  filter(year >= 1995 & city == "Minneapolis, MN") %>%
  ggplot(aes(x=fct_reorder(release, artist_familiarity, max,desc = FALSE), y=artist_familiarity)) + geom_col() + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + labs(title = "Album Releases in Minneapolis by Artist Familiarity (1995 - Present)", y = "Artist Familiarity", x = "Album Title")
arrange2
```

*The data is reordered by artist familiarity*

```{r}
arrange3 <- release_drop %>% 
  arrange(artist_familiarity) %>%
  filter(year >= 1995 & city == "Minneapolis, MN") %>%
  ggplot(aes(fct_reorder(release, artist_familiarity, max,desc = FALSE), artist_familiarity)) + geom_col() + scale_x_discrete("release") + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + labs(title = "Album Releases in Minneapolis by Artist Familiarity (1995 - Present)", y = "Artist Familiarity", x = "Album Title")
arrange3
```

*No change after arrange is used in tandem with fct_reorder*

## File I/O

*I will be looking at the top ten hottest artists during the year that I was born*

```{r}
ForCSV <- release_drop %>%
  filter(year == 1995) %>%
  group_by(artist_name) %>%
  select(release, artist_name, artist_hotttnesss) %>%
  summarise(hotness = mean(artist_hotttnesss)) %>%
  arrange(desc(hotness)) %>%
  head(10)
write_csv(ForCSV,"ForCSV.csv")
read_csv("ForCSV.csv")
```

```{r}
saveRDS(ForCSV,"ForCSV.rds")
readRDS("ForCSV.rds")
```

```{r}
dput(ForCSV,"ForCSV.txt")
dget("ForCSV.txt")
```

## Visualization Design

```{r}
viz <- singer_locations %>% 
  filter(!is.na(latitude)) %>%
  mutate(north_south=c("north", "south")[(latitude>39.43) + 1]) %>%
  filter(artist_hotttnesss != 0) %>%
  ggplot(aes(artist_hotttnesss, artist_familiarity, color = north_south)) + geom_point(alpha = .2)
viz
```

```{r}
viz2 <- singer_locations %>% 
  filter(!is.na(latitude)) %>%
  filter(latitude > 25 & latitude <= 49 & longitude < -50 & longitude > -150) %>%
  filter(artist_hotttnesss != 0) %>%
  ggplot(aes(artist_hotttnesss, artist_familiarity, color = latitude)) + 
  geom_point() +
  scale_colour_gradient2(low = "red", mid = "white", high = "blue", midpoint = 39.43, "Latitude") + 
  labs(title="Artist Hotness vs. Artist Familiarity Based on Latitude in USA", x = "Artist Hotness", y = "Artist Familiarity") 
viz2 
```

*A look at artist hotness vs artist familiarity. I used the Mason Dixon line (which divides the north from south in USA [latitude = 39.43]) as the point of divergence. I usually associated the north with a liberal agenda and the south with a conservative agenda, which informed my use of color ramp*

```{r}
singer_locations %>%
  filter(latitude > 25 & latitude <= 49 & longitude < -50 & longitude > -150) %>%
  ggplot(aes(longitude, artist_hotttnesss)) + 
  geom_point() +
  labs(title="Artist Hotness based on Latitude in USA", x = "Latitude", y = "Artist Hotness") 
```

```{r}
singer_locations %>%
  filter(latitude > 25 & latitude <= 49 & longitude < -50 & longitude > -150) %>%
  ggplot(aes(longitude, latitude, color = artist_hotttnesss)) + 
  geom_point()
```

*An attempt at looking at the distribution of artist hotness in North America*

```{r}
map <- get_googlemap(center = c(long = -97, lat = 39.8283), zoom = 4, size = c(640,640), scale = 2)
ggmap(map, fullpage = TRUE) + 
  geom_point(data=singer_locations, 
             aes(longitude, latitude, color = artist_hotttnesss)) +
  labs(title="Artist Hotness within the United States") +
  scale_colour_distiller(palette="Oranges", "Artist Hotness", direction = 1)
```

*Artist hotness ploted on a map of North America. I also tried using an intuitive color ramp.*

```{r}
mapWorld <- borders("world", colour="gray50", fill="gray50")
mp <- ggplot() + 
  mapWorld + 
  geom_point(data=singer_locations, 
             aes(longitude, latitude), colour="Deep Pink", 
             fill="Pink",pch=21, size=1, alpha=I(0.7)) +
  labs(x="Longitude", y="Latitude", title="World Map of Artist Locations")
mp 
```

*World map of artist locations. I like the contrast between pink and grey*

```{r}
a <- singer_locations %>%
  filter(!is.na(latitude)) %>%
  filter(latitude > 25 & latitude <= 49 & longitude < -50 & longitude > -150) %>%
  ggplot(aes(longitude, latitude)) + stat_density_2d(aes(fill = ..level..), geom = "polygon") + geom_point() + coord_map() 
a
```

*Here's an attempt to look at artist density within the USA. You can sort of see the shape of the USA represented by the points, but clearly a map would help with vizualization*

```{r}
usa <- map_data("usa")
POI <- singer_locations %>%
  filter(!is.na(latitude)) %>%
  filter(latitude > 25 & latitude <= 49 & longitude < -50 & longitude > -150)
densitymap <- ggplot(data = usa) + 
  geom_polygon(aes(x = long, y = lat, group = group, color = "grey"), color = "white") +  
  coord_fixed(1.3) +
  guides(fill=FALSE) +
  ##geom_point(data = POI, aes(longitude, latitude), color = "pink", alpha = 0.2) +
  stat_density_2d(data = POI, aes(longitude, latitude, color = ..level..), geom = "polygon", alpha = 0.4) + scale_colour_distiller(palette="Spectral", "Density") +
  labs(x="Longitude", y="Latitude", title="Artist Density in the United States") 
densitymap
```

*Density map of artists within the USA*

## Writing Figures to Files


```{r}
ggsave("USA_ArtistDensity.pdf", width=15, height=10, units="cm", plot = densitymap)
ggsave("ArtistDistribution.png", width=15, height=10, units="cm", plot = mp)
```

USA_ArtistDensity can be found [here](https://github.com/peterwhitman/STAT545-whitman-peter/blob/master/hw05/USA_ArtistDensity.pdf)

ArtistDistribution can be found [here](https://github.com/peterwhitman/STAT545-whitman-peter/blob/master/hw05/ArtistDistribution.png)

