---
title: "hw03"
author: "Peter Whitman"
date: "9/30/2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gapminder)
library(knitr)
library(tidyverse)
```

## Task Menu

* Get the maximum and minimum of GDP per capita for all continents. 

```{r}
gapminder %>% 
  group_by(continent) %>% filter(year == 2007) %>% summarize(MaxGDP = max(gdpPercap), MinGDP = min(gdpPercap)) 

```

*Above is a table that lists the maximum and minimum GDP/capita for each continent in 2007. I thought that focusing on a specific survey year would minimize distortions associated with things like inflation*

```{r}
p1 <- gapminder %>% 
  group_by(continent) %>% filter(year == 2007) %>% summarize(MaxGDP = max(gdpPercap), MinGDP = min(gdpPercap)) 
p2 <- p1 %>%
    group_by(continent) %>%
    summarise(MaxGDP = sum(MaxGDP), MinGDP = sum(MinGDP))
p3 <- p2 %>% gather(Legend, gdpPercapita, -continent)
ggplot(p3, aes(x = continent, fill = Legend, y = gdpPercapita)) +
    geom_col(position = "dodge") + labs(x="Continent", 
          y="GDP per Capita",
          title="The Minimum and Maximum GDP of Each Continent in 2007") 
```

*Above is a bar chart that compares the maximum and minimum GDP/capita for each continent in 2007. It took a bit of time to figure out how to get the bars to not stack, which is the default. This is where the (position = "dodge") came in.*

* How is life expectancy changing over time on different continents?

```{r}
gapminder %>% 
  group_by(continent) %>% summarize(life_change = mean(lifeExp[year == 2007] - lifeExp[year == 1952]))

```

*The above table depicts the change in life expectancy rate between 1952 and 2007 for each continent*

```{r}
p4 <- gapminder %>% 
  group_by(continent) %>% summarize(life_change = mean(lifeExp[year == 2007] - lifeExp[year == 1952])) %>% ggplot(aes(continent, life_change))
p4 + geom_col() + labs(x="Continent", 
          y="Change in Life Expectancy",
          title="Changes in Life Expectancy Between 1952 and 2007") 
```

*Above is a bar chart the depicts the data presented in the table*

```{r}
gapminder %>% group_by(continent) %>% ggplot(aes(year, lifeExp, color = continent)) + geom_smooth() + labs(x="Year", 
          y="Life Expectancy",
          title="Changes in Life Expectancy Over Time") + scale_colour_discrete("Continents of\n the World")
```

*I found that this figure depicts how life expectancy has changed over time better than total change depicted in the bar graph above*

* Report the absolute and/or relative abundance of countries with low life expectancy over time by continent: Compute some measure of worldwide life expectancy – you decide – a mean or median or some other quantile or perhaps your current age. Then determine how many countries on each continent have a life expectancy less than this benchmark, for each year.

```{r}
gapminder %>% summarize(mean(lifeExp))
```

*This is the mean life expectancy rate computed from all 1704 life expectancy entries*

```{r}
gapminder %>% mutate(benchmark=c("below", "over")[(lifeExp>59.4744) + 1]) %>% group_by(year) %>% filter(benchmark == "below") %>% count(benchmark) 
```

*This table provides the number of countries in each year that fall below the mean life expectancy rate*

```{r}
gapminder %>% mutate(benchmark=c("below", "over")[(lifeExp>59.4744) + 1]) %>% group_by(year) %>% filter(benchmark=="below") %>% count(benchmark) %>% ggplot(aes(year,n)) + geom_col() + labs(x="Year", 
          y="Number of Countries",
          title="Number of Countries Below the Mean Life Expectancy Rate Each Year")

```

*This bar chart provides a depiction of the data found in the table above*

```{r}
gapminder %>% mutate(benchmark=c("below", "over")[(lifeExp>59.4744) + 1]) %>% group_by(continent) %>% filter(benchmark == "below") %>% count(benchmark, year)
```

*Here I've computed the number of countries that fall below the mean life expectancy rate in each year by continent*

```{r}
gapminder %>% mutate(benchmark=c("below", "over")[(lifeExp>59.4744) + 1]) %>% group_by(continent) %>% filter(benchmark == "below") %>% count(benchmark, year) %>% ggplot(aes(year, n, color = continent)) + geom_smooth() + labs(x="Year", 
          y="Number of Countries Below Mean Life Expectancy Rate",
          title="Changes in Number of Countries Below Mean Life Expectancy Rate Over Time") + scale_colour_discrete("Continents of\n the World")
```

*Interestingly, Oceania is not present in this figure, which must mean that none of the countries that make up Oceania had life expectancy rates lower than the mean*