---
title: "hw10"
author: "Peter Whitman"
date: "11/30/2017"
output: github_document
---

```{r setup, include=FALSE}
library(gapminder)
library(geonames)
library(tidyverse)
library(tidyr)
library(magrittr)
library(purrr)
library(glue)
library(stringr)
library(rvest)
library(xml2)
library(ggplot2)
library(countrycode)
library(rebird)
library(rplos)
library(knitr)
```

# Use an R package that wraps an API

I followed the first prompt. Specifically, I combined gapminder and geonames data to plot population density over time. 

### 1. Make a user account and use geonames to access data about the world’s countries.

```{r}
options(geonamesUsername="pwhitman") 
```

### 2. Make a dataframe out of the country info data from geonames
```{r}
info <- GNcountryInfo()
head(info)
```

### 3. Convert the country codes within the country info data frame to the english names they represent
```{r}
info_updated <- info %>%
  mutate(country = countrycode(info$isoAlpha3, origin = "iso3c", destination = "country.name.en"))
head(info_updated)
```

### 4. Join the geonames data to gapminder data by country

```{r}
country_joined <- left_join(gapminder, info_updated, by = "country")
```

### 5. Create a new variable for population density by dividing population by area in sqkm 
```{r}
refined <- country_joined %>%
  mutate(area = as.numeric(areaInSqKm)) %>%
  select(country, continent.x, year, lifeExp, gdpPercap, pop, capital, area) %>%
  mutate(pop_density = pop/area)
kable(head(refined))
```

### 6. Vizualize the change in population density over time for each country faceted by continent using ggplot
```{r}
p <- refined %>% 
  #filter(continent.x != "Oceania") %>%
  ggplot(aes(x = year, y = pop_density, group = country, color = country)) +
  geom_line(lwd = 1, show.legend = FALSE) + 
  facet_wrap(~ continent.x) +
  labs(x = "Year", y = "Population Density (people/" ~ km^{2} ~ ")", title = "Changes in Population Density Over Time by Continent") +
  scale_color_manual(values = country_colors) + 
  theme_bw() +
  theme(strip.text = element_text(size = rel(1.1))) + 
  scale_y_log10()
p
```





