---
title: "hw01_gapminder"
author: "Peter Whitman"
date: "9/17/2017"
output: github_document
---
###1. Initialize gapminder and tidyverse
```{r}
library(gapminder)
library(tidyverse)
```
###2. Plot life expentancy by year
```{r}
plot(lifeExp ~ year, gapminder)
```

###3. Plot life expentancy by GDP per capita

```{r}
plot(lifeExp ~ gdpPercap, gapminder)
```

###4. Plot life expentancy by log(GDP per capita)

```{r}
plot(lifeExp ~ log(gdpPercap), gapminder)
```

###5. Provide a summary of the Life Expectancy data

```{r}
summary(gapminder$lifeExp)
```
###6. Produce frequency distribution histogram for Life Expectancy
```{r}
hist(gapminder$lifeExp)
```

###7. Create a frequency distribution bar plot for each continent

```{r}
barplot(table(gapminder$continent))
```

###8. Explore the relationship between GDP and Life expectancy and the variation between continents

```{r}
p <- ggplot(filter(gapminder, continent != "Oceania"),
            aes(x = gdpPercap, y = lifeExp)) # just initializes
p <- p + scale_x_log10() # log the x axis the right way
p + geom_point() # scatterplot
p + geom_point(aes(color = continent)) # map continent to color
p + geom_point(alpha = (1/3), size = 3) + geom_smooth(lwd = 3, se = FALSE)
p + geom_point(alpha = (1/3), size = 3) + facet_wrap(~ continent) +
  geom_smooth(lwd = 1.5, se = FALSE)
```


